TO FIND

### 2013-09-25 revisions/ms.tex
Tim revisions

The formulation of the recruitment probability is the central piece of all
neutral models, making the coupling with the metacommunity and
neighbouring patches possible. We adopt a simple formulation based on
\parencite{Gravel2006}. The approach is conceptually similar to placing a trap
in a canopy gap and picking a seed at random among the ones falling in to
determine the identity of the recruited species. The composition of the seed
pool in that trap will be a mixture of local dispersal and immigrants from the
metacommunity. For simplicity, we consider three spatial scales of dispersal but
it would be easy to generalize the approach to a continuous seed dispersal
kernel \parencite{Gravel2006}. The parameter $m$ is the probability that the
recruit is a migrant from neighbouring patches, $M$ is the probability it comes
from a larger (and fixed) metacommunity, and consequently, by substraction,
$1-m-M$ is the probability it comes from local dispersal. The fraction
$N_{ix}J_x^{-1}$ is the local relative abundance and $P_{ix}$ is the relative
abundance of species $i$ in the seed pool coming from patches connected to patch $x$.
The relative abundance in the neighborhood is weighted by the degree of the
connected nodes because some nodes will spread their seeds across a higher
number of nodes and thus contribute less to the seed pool. We thus consider
$P_{ix} =\frac{\sum P_{iy}d_y^{-1}}{\sum d_y^{-1}}$. We assume for simplicity
that the relative abundance in the metacommunity is uniform, i.e. equal to
$S^{-1}$, but other distributions could be used. The metacommunity is usually
modelled with a log-series (e.g. \textcite{Hubbell2001}), but it has been shown
that for a range of realistic dispersal rates a uniform distribution does not
have distinguishable impacts relative to a log-series \parencite{Bell2000}. This
immigration prevents the collapse of the metacommunity to a single species,
since otherwise all species except one will face extinction by ecological drift
(speciation prevents this phenomenon to occur in \textcite{Hubbell2001}). The
local recruitment probability $Pr_{ix}$ is consequently $Pr_{ix} = MS^{-1} +
mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. Self-replacement occurs when an individual of
species $i$ is replaced by an offspring of the same species coming either from
the remaining local individuals or from the metacommunity.

@@

The lottery dynamics described above for the neutral model assume there is a
very large number of offspring that are candidate for recruitment but only one
will survive and develop to the adult stage. The effect of a specific response to
local environmental conditions could be implemented at this stage with a biased
survival probability. The $J_x$ individuals all experience a unique
environmental condition $E_{nx}$ called a microsite $n$. We considered a patch
average $\overline{E_x}$, with a within-patch variance $\sigma_x$. The regional
average is $\overline{E_R}$ and the regional variance $\sigma_R$ (for simplicity
we considered normal distributions of environmental conditions, but different
distributions will lead to different regional similarity constraints
\parencite{Mouquet2003,Tilman2004,Gravel2006}). We consider that a fraction
$\lambda_{inx}$ of offsprings reaching the microsite where recruitment occurs
will survive. The recruitment probability is therefore biased in favour of the
species with highest survival because only some species will be able to cope
with the microsite environmental conditions. We define the relative
abundance in the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} +
(1-m-M)N_{ix}J_x^{-1}$. The calculation of the relative abundance in the seed
rain is the same as the neutral model but the recruitment probability differs
because only a fraction of offspring survive. It is formulated as $Pr_{ix} =
\frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing
the relationship between a microsite condition and survival could take various
forms; we used the traditional gaussian curve describing the niche,
$\lambda_{inx} = \exp{-\frac{(E_{nx}-u_i)^2}{2\Pi b_i^2}}$, where $u_i$ is the
niche optimum and $b_i$ is niche breadth. Note that the model will converge to a
neutral model when the niche breadth tends to infinity (which is in fact how we
simulated neutral dynamics in the Supplementary Material to minimize the
complexity of the code).

Analyses of similar models with a combination of dispersal and species sorting
shows that predictions are extremely variables and depend on the frequency
distributions of environmental conditions, niche optimums and niche breadth. For
instance, a well-studied prediction of neutral models is the species abundance
distribution. It was shown that niche models can predict similar distributions
given appropriate parameters \parencite{Tilman2004,Gravel2006}. The main
prediction is nonetheless that stable and predictable coexistence is possible if
species are sufficiently dissimilar, which differs from neutral models. Local
species richness will first depend on the joint effects of local heterogeneity
and niche breadth because coexistence requires a sufficient dissimilarity among
species \parencite{Schwilk2005}. Local species richness could be increased by a
mass effect when dispersal is consistently supplying individuals coming from
more favorable locations (refuges). The limiting similarity required to maintain
regional coexistence depends on the amount of dispersal because exchanges among
communities homogenizes environmental conditions. This is one of the main results
from the species sorting theory and a clever example of local-regional
feedbacks: increasing dispersal promotes local coexistence, but on the other
hand it diminishes regional coexistence. Only the best average competitors will
remain at very high dispersal. We therefore expect
a hump-shaped relationship between dispersal and alpha ($\alpha$) diversity,
with a peak at intermediate dispersal. On the other hand, we expect a monotonic
decrease of beta ($\beta$) and $\gamma$ diversity with increasing dispersal
\parencite{Mouquet2003}. This prediction has been validated in some experiments
\parencite{Venail2008, Logue2011}.

@@

In this section we provide simple simulation results to illustrate the impact of
spatial contingencies on species distribution and coexistence. We consider four
different landscapes, illustrated at Fig. 1. with the outcome of simulations
using the neutral model. All of these networks have the same number of nodes,
but differ both in number of edges (e.g. dispersal routes between sampling
sites) and connectivity between nodes. We ask how these differences
in topological structures will shape the emerging properties of the community under the
scenarios represented by each metacommunity model. Our analysis is not
exhaustive, it is provided simply to illustrate the interaction between
metacommunity perspectives and landscape structures on $\alpha$, $\beta$ and
$\gamma$ diversity.

@@

The network approach to the study of spatially explicit landscapes was a major
advancement in metacommunity ecology. It is a first step to make the concept
operational because it accounts for more realistic landscape structures and
dispersal kernels. It is a significant departure to island-mainland or
global dispersal approaches used previously (e.g. \parencite{Tilman1994a,
Hubbell2001, Mouquet2002}). But dispersal is also spatially explicit in a
lattice model and it does not make the landscape more realistic. We believe the
fundamental contribution of this approach is accounting for spatial
heterogeneity of dispersal. In agreement with previous theoretical
\parencite{Economo2011, Desjardins2012b} and experimemental studies
\parencite{Carrara2012}, the simulations show that the degree centrality has a
significant impact on $\alpha$ diversity. Central nodes might also be important
to maintain $\gamma$ diversity, as they promote spreading
throughout the landscape. Such nodes could be potentially quantified as keystone
for the metacommunity \parencite{Mouquet2013}. Interestingly, but not
surprisingly, this effect is weaker with species sorting dynamics. We could even
hypothesize it will vanish with very strong niche differentiation (which would occur with
low niche overlap for instance) and low mass effect. In this particular case, the
neutral versus niche comparison therefore illustrates that very strong unequal
competitive interactions could overwhelm the impact of dispersal.

@@

The multivariate variance partitioning framework originally proposed by
\textcite{Borcard1992} and further developed by \textcite{Borcard2002} has been
widely used to quantify the relative importance of species sorting and dispersal
limitations in species distribution. This framework was originally proposed to
model species distribution as a function of environmental variables, taking into
account the spatial autocorrelation of species distribution
\parencite{Leduc1992, Borcard1992, Legendre1993}. This methodology has been
widely used over the last decade as a test of the neutral theory, its underlying
assumption and to quantify dispersal limitations (e.g. \textcite{Svenning2004,
Hardy2004, Gilbert2004, Cottenie2005}. This approach is however only a weak test
of neutrality \parencite{McGill2003}, based on the description of spatial
community structure, rather than hypothesis testing. The different models we
reviewed in this article could be better employed if used to generate null
expectations of species distribution based on different hypotheses and then
compare them. But most of all, parametrized spatially explicit neutral models
could be useful if used to predict biodiversity under different global change
scenarios. For instance in freshwater systems, neutral models combined with a
dendritic network analysis were shown to predict well the distribution of
diversity in central United-States \parencite{Muneepeerakul2008}. Fitted neutral
models could be used to predict the consequences of habitat destruction,
fragmentation or a change in the connectivity matrix \parencite{Hubbell2008}.
The spatially explicit description of the landscape is a major improvement
toward that end, providing much flexibility in the scenarios that could be
explored.

Working with more complex representations on landscapes has several advantages.
Real landscapes are not flat geometric objects, they are highly structured and
diversified. This structure has long been recognized as a key component of
biological diversity. In the 19th century, Moritz Wagner noted that patterns of
rivers \cite{coy04} could explain how beetles diversified. Yet, to this day, the
relationship between spatial structures and biodiversity is not resolved, but
networks provide powerful tools to analyze landscapes and generate testable
predictions on community assembly. For example, the neutral theory predicts less
diversity in isolated communities and constant speciation (regardless of
isolation) \parencite{Economo2008}. Adding the effect of gene flow changes the
predictions \parencite{Desjardins2012b}, but the neutral theory does not predict
adaptive radiation and might thus face problems for predicting many long-term
evolutionary processes.

@@

%------------------------
\subsection*{Figure 2}
\textbf{Relationship between $\alpha$ diversity and node centrality}. The upper
two panels are simulation results conducted with the random geometric graph
illustrated at Fig. 1 and the lower two panels are runs with the random tree
graph. Parameters:$S = 100$, $c = 0.4$, $e = 0.1$, $J_x = 100$, $m = 0.2$, $M =
0.01$, $k = 0.1$, $ u \epsilon [0,100]$, $b = 15$, $ E_x \epsilon [0,100]$,
$\overline{E_R} = 50$, $\sigma{E_R} = 5$. Simulations run 1000 time steps.

>>>>

The formulation of the recruitment probability is the central piece of all
neutral models, making the coupling with the metacommunity and
neighbouring patches possible. We adopt a simple formulation based on
\parencite{Gravel2006}. The approach is conceptually similar to placing a trap
in a canopy gap and picking a seed at random among the ones falling in to
determine the identity of the recruited species. The composition of the seed
pool in that trap will be a mixture of local dispersal and immigrants from the
metacommunity. For simplicity, we consider three spatial scales of dispersal but
it would be easy to generalize the approach to a continuous seed dispersal
kernel \parencite{Gravel2006}. The parameter $m$ is the probability that the
recruit is a migrant from neighbouring patches, $M$ is the probability it comes
from a larger (and fixed) metacommunity, and consequently, by substraction,
$1-m-M$ is the probability it comes from local dispersal. The fraction
$N_{ix}J_x^{-1}$ is the local relative abundance and $P_{ix}$ is the relative
abundance of species $i$ in the seed pool coming from patches connected to patch $x$.
The relative abundance in the neighborhood is weighted by the degree of the
connected nodes because some nodes will spread their seeds across a higher
number of nodes and thus contribute less to the seed pool. We thus consider
$P_{ix} =\sum P_{iy}d_y^{-1}/\sum d_y^{-1}$. We assume for simplicity
that the relative abundance in the metacommunity is uniform, i.e. equal to
$S^{-1}$, but other distributions could be used. The metacommunity is usually
modelled with a log-series (e.g. \textcite{Hubbell2001}), but it has been shown
that for a range of realistic dispersal rates a uniform distribution does not
have distinguishable impacts relative to a log-series \parencite{Bell2000}. This
immigration prevents the collapse of the metacommunity to a single species,
since otherwise all species except one will face extinction by ecological drift
(speciation prevents this phenomenon to occur in \textcite{Hubbell2001}). The
local recruitment probability $Pr_{ix}$ is consequently $Pr_{ix} = MS^{-1} +
mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. Self-replacement occurs when an individual of
species $i$ is replaced by an offspring of the same species coming either from
the remaining local individuals or from the metacommunity.

@@

The lottery dynamics described above for the neutral model assume there is a
very large number of offspring that are candidate for recruitment but only one
will survive and develop to the adult stage. The effect of a specific response to
local environmental conditions could be implemented at this stage with a biased
survival probability. The $J_x$ individuals all experience a unique
environmental condition $E_{nx}$ called a microsite $n$. We considered a patch
average $\overline{E_x}$, with a within-patch variance $\sigma_x$. The regional
average is $\overline{E_R}$ and the regional variance $\sigma_R$ (for simplicity
we considered normal distributions of environmental conditions, but different
distributions will lead to different regional similarity constraints
\parencite{Mouquet2003,Tilman2004,Gravel2006}). We consider that a fraction
$\lambda_{inx}$ of offsprings reaching the microsite where recruitment occurs
will survive. The recruitment probability is therefore biased in favour of the
species with highest survival because only some species will be able to cope
with the microsite environmental conditions. We define the relative
abundance in the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} +
(1-m-M)N_{ix}J_x^{-1}$. The calculation of the relative abundance in the seed
rain is the same as the neutral model but the recruitment probability differs
because only a fraction of offspring survive. It is formulated as $Pr_{ix} =
\lambda_{inx}Z_{ix}/\sum \lambda_{jnx}Z_{jx}$. The function describing
the relationship between a microsite condition and survival could take various
forms; we used the traditional gaussian curve describing the niche,
$\lambda_{inx} = \exp{-\frac{(E_{nx}-u_i)^2}{2\Pi b_i^2}}$, where $u_i$ is the
niche optimum and $b_i$ is niche breadth. Note that the model will converge to a
neutral model when the niche breadth tends to infinity (which is in fact how we
simulated neutral dynamics in the Supplementary Material to minimize the
complexity of the code).

Analyses of similar models with a combination of dispersal and species sorting
shows that predictions are extremely variables and depend on the frequency
distributions of environmental conditions, niche optimums and niche breadth. For
instance, a well-studied prediction of neutral models is the species abundance
distribution. It was shown that niche models can predict similar distributions
given appropriate parameters \parencite{Tilman2004,Gravel2006}. The main
prediction is nonetheless that stable and predictable coexistence is possible if
species are sufficiently dissimilar, which differs from neutral models. Local
species richness will first depend on the joint effects of local heterogeneity
and niche breadth because coexistence requires a sufficient dissimilarity among
species \parencite{Schwilk2005}. Local species richness could be increased by a
mass effect when dispersal is consistently supplying individuals coming from
more favorable locations (refuges). The limiting similarity required to maintain
regional coexistence depends on the amount of dispersal because exchanges among
communities homogenizes environmental conditions. This is one of the main results
from the species sorting theory and a clever example of local-regional
feedbacks: increasing dispersal promotes local coexistence, but on the other
hand it diminishes regional coexistence. Only the best average competitors will
remain at very high dispersal. We therefore expect
a hump-shaped relationship between dispersal and alpha ($\alpha$) diversity,
with a peak at intermediate dispersal. On the other hand, we expect a monotonic
decrease of $\beta$ and $\gamma$ diversity with increasing dispersal
\parencite{Mouquet2003}. This prediction has been validated in some experiments
\parencite{Venail2008, Logue2011}.

@@

In this section we provide simple simulation results to illustrate the impact of
spatial contingencies on species distribution and coexistence. We consider four
different landscapes, illustrated at Fig. 1 with the outcome of simulations
using the neutral model. All of these networks have the same number of nodes,
but differ both in number of edges (e.g. dispersal routes between sampling
sites) and connectivity between nodes. We ask how these differences
in topological structures will shape the emerging properties of the community under the
scenarios represented by each metacommunity model. Our analysis is not
exhaustive, it is provided simply to illustrate the interaction between
metacommunity perspectives and landscape structures on $\alpha$, $\beta$ and
$\gamma$ diversity.

@@

The network approach to the study of spatially explicit landscapes was a major
advancement in metacommunity ecology. It is a first step to make the concept
operational because it accounts for more realistic landscape structures and
dispersal kernels. It is a significant departure to island-mainland or
global dispersal approaches used previously (e.g. \textcite{Tilman1994a,
Hubbell2001, Mouquet2002}). But dispersal is also spatially explicit in a
lattice model and it does not make the landscape more realistic. We believe the
fundamental contribution of this approach is accounting for spatial
heterogeneity of dispersal. In agreement with previous theoretical
\parencite{Economo2011, Desjardins2012b} and experimemental studies
\parencite{Carrara2012}, the simulations show that the degree centrality has a
significant impact on $\alpha$ diversity. Central nodes might also be important
to maintain $\gamma$ diversity, as they promote spreading
throughout the landscape. Such nodes could be potentially quantified as keystone
for the metacommunity \parencite{Mouquet2013}. Interestingly, but not
surprisingly, this effect is weaker with species sorting dynamics. We could even
hypothesize it will vanish with very strong niche differentiation (which would occur with
low niche overlap for instance) and low mass effect. In this particular case, the
neutral versus niche comparison therefore illustrates that very strong unequal
competitive interactions could overwhelm the impact of dispersal.

@@

The multivariate variance partitioning framework originally proposed by
\textcite{Borcard1992} and further developed by \textcite{Borcard2002} has been
widely used to quantify the relative importance of species sorting and dispersal
limitations in species distribution. This framework was originally proposed to
model species distribution as a function of environmental variables, taking into
account the spatial autocorrelation of species distribution
\parencite{Leduc1992, Borcard1992, Legendre1993}. This methodology has been
widely used over the last decade as a test of the neutral theory, its underlying
assumption and to quantify dispersal limitations (e.g. \textcite{Svenning2004,
Hardy2004, Gilbert2004, Cottenie2005}). This approach is however only a weak test
of neutrality \parencite{McGill2003}, based on the description of spatial
community structure, rather than hypothesis testing. The different models we
reviewed in this article could be better employed if used to generate null
expectations of species distribution based on different hypotheses and then
compare them. But most of all, parametrized spatially explicit neutral models
could be useful if used to predict biodiversity under different global change
scenarios. For instance in freshwater systems, neutral models combined with a
dendritic network analysis were shown to predict well the distribution of
diversity in central United-States \parencite{Muneepeerakul2008}. Fitted neutral
models could be used to predict the consequences of habitat destruction,
fragmentation or a change in the connectivity matrix \parencite{Hubbell2008}.
The spatially explicit description of the landscape is a major improvement
toward that end, providing much flexibility in the scenarios that could be
explored.

Working with more complex representations on landscapes has several advantages.
Real landscapes are not flat geometric objects, they are highly structured and
diversified. This structure has long been recognized as a key component of
biological diversity. In the 19th century, Moritz Wagner noted that patterns of
rivers \parencite{coy04} could explain how beetles diversified. Yet, to this day, the
relationship between spatial structures and biodiversity is not resolved, but
networks provide powerful tools to analyze landscapes and generate testable
predictions on community assembly. For example, the neutral theory predicts less
diversity in isolated communities and constant speciation (regardless of
isolation -- \parencite{Economo2008}. Adding the effect of gene flow changes the
predictions \parencite{Desjardins2012b}, but the neutral theory does not predict
adaptive radiation and might thus face problems for predicting many long-term
evolutionary processes.

@@

%------------------------
\subsection*{Figure 2}
\textbf{Relationship between $\alpha$ diversity and node centrality}. The upper
two panels are simulation results conducted with the random geometric graph
illustrated at Fig. 1 and the lower two panels are runs with the random tree
graph. Parameters:$S = 100$, $c = 0.4$, $e = 0.1$, $J_x = 100$, $m = 0.2$, $M =
0.01$, $k = 0.1$, $ u \epsilon [0,100]$, $b = 15$, $ E_x \epsilon [0,100]$,
$\overline{E_R} = 50$, $\sigma{E_R} = 5$. Simulations were run 1000 time steps.
